Plug-and-play dual-tree algorithm runtime analysis

### 2015-08-14 bounds-files/sec01-supporting-lemmas.tex
Rewording about the generalizability of the bound.

\begin{figure}
\begin{subfigure}[b]{0.495\textwidth}
  \begin{center}
    \begin{tikzpicture}[scale=0.8]
      \input{figures/cover-tree-points}
      \input{figures/cover-tree-s1}
    \end{tikzpicture}
  \end{center}
  \vspace*{-0.5em}
  \caption{Root node at scale $1$.}
  \label{fig:cover-tree-s1}
\end{subfigure}
\begin{subfigure}[b]{0.495\textwidth}
  \begin{center}
    \begin{tikzpicture}[scale=0.85]
      \input{figures/cover-tree-points}
      \input{figures/cover-tree-s0}
    \end{tikzpicture}
  \end{center}
  \vspace*{-0.5em}
  \caption{Nodes at scale $0$.}
  \label{fig:cover-tree-s0}
\end{subfigure}
\begin{subfigure}[b]{0.355\textwidth}
  \begin{center}
    \begin{tikzpicture}[scale=0.85]
      \input{figures/cover-tree-points}
      \input{figures/cover-tree-s-1}
    \end{tikzpicture}
  \end{center}
  \vspace*{-0.5em}
  \caption{Nodes at scale $-1$.}
  \label{fig:cover-tree-s-1}
\end{subfigure}
\begin{subfigure}[b]{0.64\textwidth}
  \begin{center}
    \begin{tikzpicture}[scale=0.6]
      \input{figures/cover-tree-ex}
    \end{tikzpicture}
  \end{center}
  \vspace*{-0.5em}
  \caption{Abstract representation.}
  \label{fig:cover-tree-ex}
\end{subfigure}
  \vspace*{-0.5em}
\caption{Example cover tree on six points.  (a) $\mathscr{N}_a$ is centered at
$p_0$ with scale $1$.  (b) $\mathscr{N}_b$ and $\mathscr{N}_c$ are centered at
$p_0$ and $p_1$, respectively, and have scale $0$.  (c) $\mathscr{N}_d$ and
$\mathscr{N}_e$ are centered at $p_0$ and $p_2$, respectively, and have scale
$-1$.  The leaves, $\mathscr{N}_0$ through $\mathscr{N}_6$, are centered at each
of the six points, with scale $-\infty$ (and therefore radius 0).  Note that
although node $\mathscr{N}_b$ in subfigure (b) overlaps node $\mathscr{N}_c$,
point $p_1$ only belongs to $\mathscr{N}_c$, not $\mathscr{N}_b$.  Note also
that this is only one valid cover tree that could be built on the data; other
configurations are possible; for instance, selecting a different root point
gives different valid cover trees.}
\label{fig:example-cover-tree}
\end{figure}

@@

The expansion constant is used heavily in the cover tree literature.  It is,
in some sense, a notion of instrinic dimensionality, and previous work has shown
that there are many scenarios where $c$ is independent of the number of points
in the dataset \citep{karger2002finding, langford2006,
krauthgamer2004navigating, ram2009}.  Note also that if points in $S \subset
\mathcal{H}$ are being drawn according to a stationary distribution $f(x)$, then
$c$ will converge to some finite value $c_f$ as $|S| \to \infty$.  To see this,
define $c_f$ as a generalization of the expansion constant for distributions.
$c_f \ge 2$ is the smallest value such that

>>>>

\begin{figure}
\begin{subfigure}[b]{0.495\textwidth}
  \begin{center}
    \begin{tikzpicture}[scale=0.8]
      \input{figures/cover-tree-points}
      \input{figures/cover-tree-s1}
    \end{tikzpicture}
  \end{center}
  \vspace*{-0.5em}
  \caption{Root node at scale $1$.}
  \label{fig:cover-tree-s1}
\end{subfigure}
\begin{subfigure}[b]{0.495\textwidth}
  \begin{center}
    \begin{tikzpicture}[scale=0.85]
      \input{figures/cover-tree-points}
      \input{figures/cover-tree-s0}
    \end{tikzpicture}
  \end{center}
  \vspace*{-0.5em}
  \caption{Nodes at scale $0$.}
  \label{fig:cover-tree-s0}
\end{subfigure}
\begin{subfigure}[b]{0.355\textwidth}
  \begin{center}
    \begin{tikzpicture}[scale=0.85]
      \input{figures/cover-tree-points}
      \input{figures/cover-tree-s-1}
    \end{tikzpicture}
  \end{center}
  \vspace*{-0.5em}
  \caption{Nodes at scale $-1$.}
  \label{fig:cover-tree-s-1}
\end{subfigure}
\begin{subfigure}[b]{0.64\textwidth}
  \begin{center}
    \begin{tikzpicture}[scale=0.6]
      \input{figures/cover-tree-ex}
    \end{tikzpicture}
  \end{center}
  \vspace*{-0.5em}
  \caption{Abstract representation.}
  \label{fig:cover-tree-ex}
\end{subfigure}
  \vspace*{-0.5em}
\caption{Example cover tree on six points in $\mathcal{R}^2$.  (a)
$\mathscr{N}_a$ is centered at $p_0$ with scale $1$.  (b) $\mathscr{N}_b$ and
$\mathscr{N}_c$ are centered at $p_0$ and $p_1$, respectively, and have scale
$0$.  (c) $\mathscr{N}_d$ and $\mathscr{N}_e$ are centered at $p_0$ and $p_2$,
respectively, and have scale $-1$.  The leaves, $\mathscr{N}_0$ through
$\mathscr{N}_6$, are centered at each of the six points, with scale $-\infty$
(and therefore radius 0).  Note that although node $\mathscr{N}_b$ in subfigure
(b) overlaps node $\mathscr{N}_c$, point $p_1$ only belongs to $\mathscr{N}_c$,
not $\mathscr{N}_b$.  Note also that this is only one valid cover tree that
could be built on the data; other configurations are possible; for instance,
selecting a different root point gives different valid cover trees.}
\label{fig:example-cover-tree}
\end{figure}

@@

The expansion constant is used heavily in the cover tree literature.  It is,
in some sense, a notion of instrinic dimensionality, most useful in scenarios
where $c$ is independent of the number of points in the dataset
\citep{karger2002finding, langford2006, krauthgamer2004navigating, ram2009}.
Note also that if points in $S \subset \mathcal{H}$ are being drawn according to
a stationary distribution $f(x)$, then $c$ will converge to some finite value
$c_f$ as $|S| \to \infty$.  To see this, define $c_f$ as a generalization of the
expansion constant for distributions.  $c_f \ge 2$ is the smallest value such
that

### 2015-08-14 bounds.tex
Add an example cover tree figure.
Rewording about the generalizability of the bound.

@@

Section \ref{sec:preliminaries} lays out the required background, notation, and
introduces the cover tree and its associated theoretical properties.  Readers
familiar with the cover tree literature and dual-tree algorithms
\citep[especially][]{curtin2013tree} may find that section to be review.
Following that, we introduce an intuitive measure of cover tree imbalance, an
important property for understanding the runtime of dual-tree algorithms, in
Section \ref{sec:imbalance}.  This measure of imbalance is then used to prove
the main result of the paper in Section \ref{sec:bound}, which is a worst-case
runtime bound for generalized dual-tree algorithms.  We apply this result to
three specific problems: nearest neighbor search (Section \ref{sec:nns}),
approximate kernel density estimation (Section \ref{sec:akde}), and range search
/ range count (Section \ref{sec:rs}), showing linear runtime bounds for each of
those algorithms.  Each of these bounds is an improvement on the
state-of-the-art, and in the case of range search, is the first such bound.
%  We
%then investigate the empirical performance of cover trees for these algorithms
%in Section \ref{sec:emp}, with a focus on understanding the dataset-dependent
%constants that our bounds use.

>>>>

\usetikzlibrary{calc}

@@

Section \ref{sec:preliminaries} lays out the required background, notation, and
introduces the cover tree and its associated theoretical properties.  Readers
familiar with the cover tree literature and dual-tree algorithms
\citep[especially][]{curtin2013tree} may find that section to be review.
Following that, we introduce an intuitive measure of cover tree imbalance, an
important property for understanding the runtime of dual-tree algorithms, in
Section \ref{sec:imbalance}.  This measure of imbalance is then used to prove
the main result of the paper in Section \ref{sec:bound}, which is a worst-case
runtime bound for generalized dual-tree algorithms.  We apply this result to
three specific problems: nearest neighbor search (Section \ref{sec:nns}),
approximate kernel density estimation (Section \ref{sec:akde}), and range search
/ range count (Section \ref{sec:rs}), showing linear runtime bounds for each of
those algorithms.  Each of these bounds is an improvement on the
state-of-the-art, and in the case of range search, is the first such bound.
Despite the intuition this provides for the scaling properties of all dual-tree
algorithms\footnote{Dual-tree algorithms using $kd$-trees and other types of
trees have been observed to empirically scale linearly for tasks that take
quadratic time without the use of trees; see the empirical results of
\citet{nbody, march2010euclidean, vladymyrov2014linear, klaas2006fast,
gray2003nonparametric}.}, it must be kept in mind that these worst-case bounds
only apply to dual-tree algorithms that use the cover tree and the standard
cover tree traversal.
%  We
%then investigate the empirical performance of cover trees for these algorithms
%in Section \ref{sec:emp}, with a focus on understanding the dataset-dependent
%constants that our bounds use.
